---
title: "assignment 1"
author: "Fangzhou Yang"
date: "10/06/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
**For loops: section 4.1.1, exercises 2**

#(a)
```{r}
Fibonacci <- numeric(31)
Fibonacci[1] <- Fibonacci[2] <- 1
for (i in 3:31) Fibonacci[i] <- Fibonacci[i - 2] + Fibonacci[i - 1]

ratio<-numeric(30)
for (i in 1:30) ratio[i] <- Fibonacci[i +1] / Fibonacci[i]
cat("The sequence is:",ratio)
```

Yes, the sequence Seems converge to 1.618034.


#(b)
```{r}
golden_ratio = (1+sqrt(5))/2
print(golden_ratio)

```

Yes, the sequence is converging to the golden ratio.

#Proof
The golden ratio function is shown as:
$$1+1/\phi=\phi$$
in part (a), we get the Fibonacci sequence f<sub>n</sub>, and calculate the ratio of f<sub>n</sub>/f<sub>n-1</sub>. Since for a Fibonacci sequence, f<sub>n</sub> = f<sub>n-1</sub> + f<sub>n-2</sub>,
So the ratio is:
$$r_n =f_n/f_{n-1}=(f_{n-1}+f_{n-2})/f_{n-1}=1+f_{n-2}/f_{n-1}=1+1/r_{n-1}$$
$$r_n-\phi=1+1/r_{n-1}-1-1/\phi=1/r_{n-1}-1/\phi=(\phi-r_{n-1})/(r_{n-1}\phi)$$
So as n goes to infinity, the value (r<sub>n</sub> -  ¦Õ) goes to 0.

Therefore the ratio is converging to  ¦Õ, which equals $$(\sqrt{5}+1)/2 = 1.618034$$.

## Question 2
**For loops: section 4.1.1, exercises 3**

(a).answer=15
```{r}
answer <- 0
for (j in 1:5) answer <- answer + j
answer
```

(b).answer={1,2,3,4,5}
```{r}
answer <- NULL
for (j in 1:5) answer <- c(answer, j)
answer
```

(c).answer={0,1,2,3,4,5}
```{r}
answer <- 0
for (j in 1:5) answer <- c(answer, j)
answer
```

(d).answer=120
```{r}
answer <- 1
for (j in 1:5) answer <- answer * j
answer
```

(e).answer={3,21,23,6,11,14,5,4,28,10,8,25,20,17,26,27}
```{r}
answer <- 3
for (j in 1:15) answer <- c(answer, (7 * answer[j]) %% 31)
answer
```

So if we rewrite the code by use j from 1 to 30(or larger), we can find that the sequence is repeated every 16 elements. So from the 17th element of the sequence, it repeats {3,21,23,5,11,15,12,22,30,24,13,29,17,26,27} again.
```{r}
answer <- 3
for (j in 1:30) answer <- c(answer, (7 * answer[j]) %% 31)
answer
```

## Question 3
**If statements: section 4.1.2, exercises 4**
```{r}
earn<-function(P,x){
  if(x<=3){
    return(P*((1.04^x)-1))
  }
  if(x>3){
    return(P*((1.05^x)-1))
  }
}
```

## Question 4
**If statements: section 4.1.2, exercises 5**

```{r}
mortgage<-function(n,P,open){
  if(open==TRUE){
    return((0.005*P)/(1-((1.005)^(-n))))
  }
  else{
    return((0.004*P)/(1-((1.004)^(-n))))
  }
}
```

## Question 5
**while statements: section 4.1.3, exercises 2**

```{r}
Fibonacci <- c(1, 1)
#Fibonacci[length(Fibonacci)]
#Fibonacci[length(Fibonacci)-1]
#Fibonacci[length(Fibonacci)-1] + Fibonacci[length(Fibonacci)]

while ((Fibonacci[length(Fibonacci)-1] + Fibonacci[length(Fibonacci)]) < 300) 
{
  Fibonacci <- c(Fibonacci, (Fibonacci[length(Fibonacci)-1] + Fibonacci[length(Fibonacci)]))
}
Fibonacci

```

## Question 6
**while statements: section 4.1.3, exercises 4**
```{r}
i <- 0.006
diff <- 1

while (diff > 0.000001)
{
  diff <- abs(i-((1-((1+i)^(-20)))/19))
  i <- (1-((1+i)^(-20)))/19
}
print(i)

```
I have tried other starting guesses(which shows as below), and can find that the value we get are almost the same, which is around 0.0049~0.00495.

```{r}
i <- 0.002
diff <- 1

while (diff > 0.000001)
{
  diff <- abs(i-((1-((1+i)^(-20)))/19))
  i <- (1-((1+i)^(-20)))/19
}
print(i)
```

## Question 7
**while statements: section 4.1.3, exercises 5**

```{r}
i <- 0.006
diff <- 1
n <- 1
while (diff > 0.000001)
{
  diff <- abs(i-((1-((1+i)^(-20)))/19))
  i <- (1-((1+i)^(-20)))/19
  n <- n+1
}
print(i)
cat("the number of iterations required to get two successive values of the interest rate that are less than 0.000001 apart is",n,".")

```

## Question 8
**break statements: section 4.1.5, exercises 2**

# (a)
```{r}
Eratosthenes <- function(n) {
# Print prime numbers up to n (based on the sieve of Eratosthenes)
if (n >= 2) {
 sieve <- seq(2, n)
 primes <- c()
while (length(sieve) > 0) 
  {
 p <- sieve[1]
 primes <- c(primes, p)
 sieve <- sieve[(sieve %% p) != 0]
  }
return(primes)
  } else {
 stop("Input value of n should be at least 2.")
  }
}
```
We'd like to find all prime numbers up to the given value n through this function. Initially, all
integers from 2 through n are stored and the primes object is set up initially empty. The composite numbers in sieve are removed, and the primes are copied to primes. We can use sieve[(sieve %% p) != 0] to eliminate all multiples of p. This step also eliminates p, but we have already saved it in primes.

# (b)
Since the elements remaining are primes, which are not multiples of integers greater than 2 and smaller than itself. And an integer(<n) which is not a multiple of any integer >= 2 and <= sqrt(n) must be a prime. 
Thus, once p >= sqrt(n), all remaining entries in sieve are prime.

# (c)
```{r}
Eratosthenes <- function(n) {
  # Print prime numbers up to n (based on the sieve of Eratosthenes)
if (n >= 2) 
  {
   sieve <- seq(2, n)
   primes <- c()
  repeat
    {
     p <- sieve[1]
     primes <- c(primes, p)
     sieve <- sieve[(sieve %% p) != 0]
     if (length(sieve) <= 0) break
    }
  return(primes)
  } else {
    stop("Input value of n should be at least 2.")
  }
}

```

## Question 9
**Functions: section 4.2.1, exercises 2**

#(a)
```{r}
compound.interest<-function(P,i.r,n){
  return(P*((1+i.r)^n))
}
```

#(b)
```{r}
compound.interest(1000,0.01,30)
```

## Question 10
**Functions: section 4.2.1, exercises 3**

```{r}
bisection <- function(eq, a, b){
 repeat{
 eq1 <- f(a)
 eq2 <- f(b)
 m <- (a+b)/2
 eq3 <- f(m)
 
 if(abs(a-b) < 0.000001){
 print((a+b)/2)
 break
 }
 
 else{
 if(eq3*eq1 > 0){
 a <- m
 }else{
 b <- m
 }
 }
 
 if(eq3 == 0){
 print(m)
 break
 }
 }
 }
```

## Question 11
**Functions: section 4.4.1, exercises 1**

```{r}
mergesort <- function (x) {
  # Check for a vector that doesn¡¯t need sorting
  len <-length(x)
  if (len < 2) result <- x
  else {
    #  sort x into result
    #  split x in half
    y <- x[1:(len / 2)]
    z <- x[(len/2+1):len]
    #  sort y and z
    y <- mergesort(y)
    z <- mergesort(z)
    #  merge y and z into a sorted result
    result <- c()
    #  while (some are left in both piles)
    while (min(length(y), length(z)) > 0) {
    #  put the largest first element on the end
    #  remove it from y or z
      if (y[1] > z[1]) {
        result <- c(result, y[1])
        y <- y[-1]
      } else {
        result <- c(result, z[1])
        z <- z[-1]
      }
    }
    #  put the leftovers onto the end of result
    if (length(y) > 0)
    result <- c(result, y)
    else
    result <- c(result, z)
  }
  return(result)
}
```

## Question 12
**Functions: section 4.4.1, exercises 2**

#(a)
```{r}
equation_2<-function(f,g){
 x<-1
  y<-1
  f_n<-eval(f)
  g_n<-eval(g)
  
  while((abs(f_n)>0.0000001)|(abs(g_n)>0.0000001)){
    f_x<-eval(D(f,"x"))
    g_x<-eval(D(g,"x"))
    f_y<-eval(D(f,"y"))
    g_y<-eval(D(g,"y"))
    
    d_n <- f_x*g_y - g_x*f_y
    x <- x-(g_y*f_n - f_y*g_n)/d_n
    y <- y-(f_x*g_n - g_x*f_n)/d_n
    
    f_n<-eval(f)
    g_n<-eval(g)
  }
  print(c(x,y))
}

```


#(b)
```{r}
equation_2(expression(x+y),expression(x^2+2*(y^2)-2))
```

So through the process, we can get x=-0.8164966, and y=0.8164966. If we solve this question, we should get x=sqrt(2/3), y=-sqrt(2/3) or x=-sqrt(2/3), y=sqrt(2/3). We can see that 0.8164966 is pretty close to sqrt(2/3), which check the result. 

## Question 13
**Chapter 4 exercises 1**

```{r}
directpoly<-function(x,c){
  P<-0
  n<-length(c)
  for(i in 1:n){
    P<-P+c[i]*(x^(n-i))
  }
  return(P)
}
```

## Question 14
**Chapter 4 exercises 2**

```{r}
hornerpoly<-function(x,c){
  a<- matrix(0,length(x),length(c))
  for(i in (length(c)-1):1){
    a[,i] <- a[,i+1]*x+c[i]
  }
  return(a[,1])
}

```

## Question 15
**Chapter 4 exercises 3**

#(a)
```{r}
directpoly<-function(x,c){
  P<-0
  n<-length(c)
  for(i in 1:n){
    P<-P+c[i]*(x^(n-i))
  }
  return(P)
}
hornerpoly<-function(x,c){
  a<- matrix(0,length(x),length(c))
  for(i in (length(c)-1):1){
    a[,i] <- a[,i+1]*x+c[i]
  }
  return(a[,1])
}
system.time(directpoly(x=seq(-10, 10, length=5000000), c(1, -2, 2, 3, 4, 6, 7)))
system.time(hornerpoly(x=seq(-10, 10, length=5000000), c(1, -2, 2, 3, 4, 6, 7)))

```

From the process, we can see that the hornerpoly method is much quicker than the directpoly. The system processing time of directpoly is 1.92, 0.10, 2.03, and time of hornorpoly is 0.74,0.22,0.96 which is much quicker.

#(b)
```{r}
system.time(directpoly(x=seq(-10, 10, length=5000000), c(2, 17, -3)))
system.time(hornerpoly(x=seq(-10, 10, length=5000000), c(2, 17, -3)))
```


Both method could process quickly when the number of polynomial coefficients is smaller. But the system processing time of hornoerpoly method is still faster than the directpoly method.
