---
title: "Chapter 4 exercise 3"
author: "Fangzhou Yang"
date: "10/06/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 15
**Chapter 4 exercises 3**

#(a).

```{r}
directpoly<-function(x,c){
  P<-0
  n<-length(c)
  for(i in 1:n){
    P<-P+c[i]*(x^(n-i))
  }
  return(P)
}
hornerpoly<-function(x,c){
  a<- matrix(0,length(x),length(c))
  for(i in (length(c)-1):1){
    a[,i] <- a[,i+1]*x+c[i]
  }
  return(a[,1])
}
system.time(directpoly(x=seq(-10, 10, length=5000000), c(1, -2, 2, 3, 4, 6, 7)))
system.time(hornerpoly(x=seq(-10, 10, length=5000000), c(1, -2, 2, 3, 4, 6, 7)))

```

From the process, we can see that the hornerpoly method is much quicker than the directpoly. The system processing time of directpoly is 1.64, 0.08, 1.75, and time of hornorpoly is 0.66,0.17,0.84 which is much quicker.

#(b).
```{r}
system.time(directpoly(x=seq(-10, 10, length=5000000), c(2, 17, -3)))
system.time(hornerpoly(x=seq(-10, 10, length=5000000), c(2, 17, -3)))
```


Both method could process quickly when the number of polynomial coefficients is smaller. But the system processing time of hornoerpoly method is still faster than the directpoly method.