---
title: 'For loops exercise 2'
author: "Fangzhou Yang"
date: "10/03/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
**For loops: section 4.1.1, exercises 2**

#(a)
```{r}
Fibonacci <- numeric(31)
Fibonacci[1] <- Fibonacci[2] <- 1
for (i in 3:31) Fibonacci[i] <- Fibonacci[i - 2] + Fibonacci[i - 1]

ratio<-numeric(30)
for (i in 1:30) ratio[i] <- Fibonacci[i +1] / Fibonacci[i]
cat("The sequence is:",ratio)
```

Yes, the sequence Seems converge to 1.618034.


#(b)
```{r}
golden_ratio = (1+sqrt(5))/2
print(golden_ratio)

```

Yes, the sequence is converging to the golden ratio.

#Proof
The golden ratio function is shown as:
$$1+1/\phi=\phi$$
in part (a), we get the Fibonacci sequence f<sub>n</sub>, and calculate the ratio of f<sub>n</sub>/f<sub>n-1</sub>. Since for a Fibonacci sequence, f<sub>n</sub> = f<sub>n-1</sub> + f<sub>n-2</sub>,
So the ratio is:
$$r_n =f_n/f_{n-1}=(f_{n-1}+f_{n-2})/f_{n-1}=1+f_{n-2}/f_{n-1}=1+1/r_{n-1}$$
$$r_n-\phi=1+1/r_{n-1}-1-1/\phi=1/r_{n-1}-1/\phi=(\phi-r_{n-1})/(r_{n-1}\phi)$$
So as n goes to infinity, the value (r<sub>n</sub> -  ¦Õ) goes to 0.

Therefore the ratio is converging to  ¦Õ, which equals $$(\sqrt{5}+1)/2 = 1.618034$$.